---
title: ""
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle_thesis.docx
  pdf_document: default
  html_document:
    number_sections: yes
---

```{r, results="asis", echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

\
**Resource allocation in *Eucalyptus***
\
\
\
By

\
**Courtney Campany**
\
\
\
**A thesis submitted in fulfilment of the requirements**
\
**for the degree of Doctor of Philosophy**
\

```{r concept, echo=FALSE, error=TRUE,fig.width=3, fig.height=2,dev="png", dpi=600, dev.args=list(type="cairo")}
library(png)
concept <- readPNG("thesis/WSU_HIE_Logo3.png")
par(mar=c(0,0,0,0))
plot(c(0,1),c(0,1), type="n", axes=F, xlab="", ylab="")
rasterImage(concept, xleft = 0, ybottom = 0, xright = 1, ytop =1)

```

\
**2016**

# Acknowledgements

“I am the Lorax. I speak for the trees. I speak for the trees for the trees have no tongues.”
--Dr. Seuss

\
I dedicate this thesis to my mother, for she has been my inspiration to become a scientist for as long as I can remember. Her passion for the natural world is infectious and her ability to pass her scientific knowledge to others is inspirational. It is because of you that I flip over rocks to see what is hiding underneath.

\
I would equally like to acknowledge my father for instilling in me the work ethic and drive that has carried me to this point. Without his undying support I might never have achieved the level of success I have been afforded.

\
À Caroline, rien de tout ça n’aurait été possible sans que tu sois à mes côtés. Tu as été un roc quand j’avais besoin d’ancrage et une lumière quand mon chemin semblait perdu.”

## Statement of Authentication
\
\
\
\
The work presented in this thesis is, to the best of my knowledge and belief, original
except as acknowledged in the text. I hereby declare that I have not submitted this
material, either in full or in part, for a degree at this or any other institution.

# Preface

# Table of Contents

# LIST OF TABLES

**Table 2.1**. Responses of plant and leaf characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments.

\
**Table 2.2**. Responses of root characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. 

\
**Table 2.3**. Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* seedlings to soil volume treatments.

\
**Table 2.S1**. Seedling growth model default parameters.

\
**Table 3.1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments.

\
**Table 3.2**. *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments.

\
**Table 4.1**. Final harvest C mass of above and belowground tissues, cumulative aboveground tree C uptake ($\text{F}_{c,T}$) and specific leaf area (SLA).

# LIST OF FIGURES

**Figure 2.1**. Soil volume treatment means &pm; standard error of height growth (a), diameter growth (b), and seedling leaf area estimated from leaf counts (c) measured weekly of *Eucalyptus tereticornis* seedlings across the experiment duration in 2013.

\
**Figure 2.2**. Daily maximum and minimum temperature (a), total daily PPFD (b), and daily maximum vapour pressure deficit (c) across the experiment duration in 2013.

\
**Figure 2.3**. Soil volume treatment means of biomass partitioning to leaves, stems, and roots at harvest (a), bi-variate relationships between mass allocation to leaves and stems + roots (b) and leaf mass as a function of fine root biomass with &pm; standard error (c).

\
**Figure 2.4**. Soil volume treatment means &pm; standard error, across all measurement campaigns (n = 6), of light saturated rates of photosynthesis at 25&deg;C.
\
**Figure 2.5**. Photosynthetic capacity, on a leaf mass basis, as a function of accumulation of leaf starch (a) and leaf nitrogen content without TNC (b).

\
**Figure 2.6**. Total carbon mass for harvested and modeled seedlings versus predicted total carbon gain after 120 days (a) and  reductions in final seedling carbon mass, both modeled and observed, as a function of the reduction in leaf photosynthesis across treatments (b).

\
**Figure 2.S1**. Sensitivity testing of seedling growth model to different carbon allocation strategies including; constraints of leaf mass fraction to treatment specific final harvest values (a) and increases in respiration of non-leaf tissue components by 50 % (b).

\
**Figure 3.1**. Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014.

\
**Figure 3.2**. (a) AC~i~ curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments.

\
**Figure 3.3**. The response of *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments. 

\
**Figure 3.4**. The mean &pm; 1 standard error of g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.

\
**Figure 3.5**. (a) Response of instantaneous transpiration efficiency (ITE) to VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.

\
**Figure 3.6**. Relationship between the observed discrimination of ^13^CO~2~ measured during photosynthesis ($\Delta_o$) and measured C~i~/C~a~ for sun leaves measured at high light and shade leaves measured at both low and high light.

\
**Figure 3.7**. The mean &pm; 1 standard error of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).

\
**Figure 3.S1**. Daily maximum and minimum temperature (a), daily maximum VPD (b) and total daily PPFD (c) for each chamber across the experiment duration.

\
**Figure 3.S2**. Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments.

\
**Figure 3.S3**. Response of *A~n~* (a), g~m~ (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.

\
**Figure 3.S4**. Response of VPD (a), g~s~ (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.

\
**Figure 4.1**. Conceptual diagram depicting the major components of C flow among plant components including; uptake via photosynthesis, allocation to component tissues, tissue respiration and root exudation.

\
**Figure 4.2**. Whole tree C mass as a function of cumulative aboveground C flux for each WTC tree. Values of cumulative aboveground net C flux were measured over the final eleven months of the experiment.

\
**Figure 4.3**. Estimated canopy leaf area for each WTC tree over the final eleven months of the experiment (April 2008 to March 2009).

\
**Figure 4.4**. Treatment means of cumulative aboveground C flux as a function of mean daily canopy leaf area over the final eleven months of the experiment.

\
**Figure 4.5**. Treatment means of C mass fractions of leaves (a), stems (branches+boles) (c) and roots (e) as a function of tree size, via whole tree C mass.

\
**Figure 4.6**. Cumulative aboveground net C flux and additive C allocation to individual tree components from 15 April 2008 to 16 March 2009.

\
**Figure 4.7**. Treatment means &pm; 1 standard error of cumulative aboveground net C flux, TBCA, and the residual belowground C flux ($\text{F}_{c,r}$).

\
**Figure 4.8**. Total belowground C allocation as a function of cumulative aboveground net C flux across the final eleven months of the experiment.
\
**Figure 4.S1**.
Root mass as a function of shoot mass in *Eucalyptus saligna* for potted seedlings harvested before planting of WTC trees (n=17) and WTC trees harvested after 2 years (n=12).

\
**Figure 4.S2**.
Cumulative aboveground net C flux and additive C allocation of individual tree components from 2008-4-15 and 2009-3-16. Panels represent each individual WTC.


# LIST OF ABBREVIATIONS

*A~n~*  Net leaf photosynthesis rate
\
aC~a~ Ambient CO~2~ treatment
\
AC~i~ Photosynthetic CO~2~ response curves 
\
A~max~ Leaf net photosynthesis at saturating light and CO~2~ concentration
\
A~sat~ Leaf net photosynthesis at saturating light
\
AT  Ambient air temperature treatment
\
C Carbon
\
[CO~2~]   CO~2~ concentration
\
C~a~  Atmospheric CO~2~ concentration
\
$\text{C}_{ab}$ Aboveground standing crop C mass
\
C~day~ Predicted daily carbon assimilation
\
C~i~ Intercellular CO~2~ concentration (or partial pressure)
\
C~c~ Chloroplastic CO~2~ concentration (or partial pressure)
\
$\text{C}_{r,T}$  Total C mass of roots
\
E Leaf transpiration
\
eC~a~ Elevated CO~2~ treatment
\
ET  Elevated air temperature treatment
\
FACE Free-air CO~2~ enrichment experiments
\
$\text{F}_{c}$  Net aboveground carbon uptake
\
$\text{F}_{c,r}$  Residual belowground C flux
\
free  freely-rooted 
\
FRLD Fine root length density
\
g~s~  Stomatal conductance
\
g~m~  Mesophyll conductance
\
ITE Leaf level instantaneous transpiration efficiency
\
J~max~ Maximum rate of photosynthetic electron transport 
\
K~l~ Leaf-specific hydraulic conductance 
\
LA Leaf area
\
LMA Leaf mass per unit area
\
LMF Leaf mass fraction
\
N  Nitrogen
\
N~a~ Leaf nitrogen on an area basis
\
N~f~ TNC-free leaf nitrogen content
\
PPFD  Photosynthetic photon flux density
\
Q~10~ Rate of change in respiration due to 10 &deg;C increase in temperature
\
*R* Leaf dark respiration rate
\
RMF Root mass fraction
\
SLA Specific leaf area
\
SLA~f~  TNC-free Specific leaf area
\
SMF Stem mass fraction
\
TBCA Total belowground C allocation
\
TNC Total non-structural carbohydrate
\
TDL Tunable diode laser
\
Vc~max~ Maximum rate of rubisco carboxylation 
\
VPD Vapour pressure deficit
\
VPDB Standard Vienna Pee Dee Belemnite
\
WUE Water-use efficiency
\
WTC Whole-tree chambers
\
$\delta$^13^C   C isotope ratios of 13^C to 12^C expressed relative to standard Vienna Pee Dee Belemnite
\
$\Delta$ C isotope discrimination during C3 photosynthesis 
\
$\Psi$~l~ Midday leaf water potential
\
$\Psi$~pd~  Predawn leaf water potential
\
$\sigma$~s~ Self shading parameter

#Abstract 

Plants must utilize external resources including light, CO~2~, water and mineral nutrients to support photosynthetic carbon (C) gain.  This photoassimilate is then allocated within the plant as the essential C resource for growth, maintenance and storage. Theory and observations suggest that C allocation and leaf physiology are optimized to maintain functional balance for external resource capture and to maximize C gain. However, the impacts of a changing climate may disrupt the proposed balance of C allocation between above and belowground pools. Variation in resource distribution and leaf physiology within tree canopies is also not fully understood, thus all canopy leaves may not follow theories of leaf optimal behavior. This lack of understanding regarding C uptake and the fate of assimilated C inhibits our ability to precisely test the coordination between canopy photosynthesis and growth. To address these broad ecological questions, this PhD research utilized a diverse set of experiments which manipulated resource availability and climate factors on *Eucalyptus* tree species. My goal was to measure aspects of resource allocation and C uptake across different scales, from leaf to whole tree, to improve understanding of the physiological processes which determine tree growth and the sensitivity of these processes to changing environments.

\
First, I determined whether manipulations of soil volume would limit growth in *Eucalyptus tereticornis* seedlings by disrupting the balance between source and sink activity. Seedlings were grown in a large range of container sizes and planted flush to the soil alongside naturally planted freely-rooted seedlings (free). Aboveground growth of seedlings in containers was negatively affected compared to free seedlings soon after the experiment started. Despite large reductions in total growth across soil volume treatments, relative partitioning of mass to leaves, stems and roots was similar for all seedlings after 120 days. Leaf photosynthetic capacity decreased in containers compared to free seedlings, and was correlated to both leaf nitrogen (N) content and starch accumulation. Although belowground sink limitation resulted in a reduction of net leaf photosynthesis (*A~n~*), a mass balance model concluded that these reductions were not large enough to explain observed growth responses. As *A~n~* and growth were not tightly coordinated, the model predicted excess photosynthetic C not attributed to biomass in potted seedlings. Quantifying the fate of this excess C will be essential in evaluating feed-backs between sink strength and leaf C uptake in future studies.

\
Second, I investigated how light gradients within *Eucalyptus tereticornis* tree canopies affect the distribution of resources that define photosynthetic capacity of sun and shade leaves. Trees were grown in climate-controlled  whole tree chambers under prevailing ambient and warmed (+3 °C) treatments and leaf gas exchange was coupled with online C isotope discrimination to measure *A~n~*, stomatal conductance (g~s~) and mesophyll conductance (g~m~) of sun and shade leaves. Photosynthesis rates were reduced by ca. 40 % in shade leaves associated with a 75 % reduction in photosynthetically active radiation compared to sun leaves. Photosynthetic capacity (ca. 20 % lower Vc~max~ and J~max~) and leaf N were also lower in shade leaves than sun leaves however; g~s~ was similar. Leaf C~i~, estimated from both leaf $\delta$^13^C and gas exchange, was higher in shade leaves than sun leaves. Here, the optimization theory hypothesis that C~i~ should be optimized throughout the canopy was rejected because water use efficiency was lower in shade leaves, compared to sun leaves. When light intensity was increased from low light to high light for shade leaves, both g~s~ and g~m~ increased rapidly (within minutes), leading to increases in *A~n~* greater than sun leaves at the same high light environment. This rapid response of g~m~ with light likely enables shade leaves to respond quickly to sunflecks and possibly represents a new mechanism underpinning leaf gas exchange responses to light. This capacity of shade leaves to adjust their physiological behavior and increase C uptake when sunflecks occur likely plays significant role in whole tree C uptake for some tree species, albeit at the cost or reduced water-use efficiency. These findings reveal that plant resources within a canopy may be distributed to utilize sunflecks and the dynamic physiological responses of shade leaves to altered light environments must be accounted for when up-scaling leaf level measurements to predict whole canopy C gain. 

\
Finally, I examined how net aboveground C uptake correlated to tree biomass growth and whether elevated [CO~2~] and drought treatments altered C allocation patterns above or belowground in *Eucalyptus saligna* trees. Trees were grown in climate-controlled whole tree chambers (WTCs) over a period of two years under treatments of two [CO~2~] (380 ppm and 620 ppm) and two watering regimes (well-watered and a four-month drought) in factorial combination. Additionally, we utilized a novel approach to calculate total belowground C allocation (TBCA) for each WTC as the residual between the measured aboveground net CO~2~ uptake and aboveground C mass. Measured cumulative aboveground net C uptake correlated positively to whole tree C mass production and leaf area over the final eleven months of the experiment. Contrary to previous studies, cumulative TBCA was unaffected by either elevated CO~2~ or drought treatments. As a fraction of total aboveground net C uptake, TBCA was also found to remain relatively stable across daily time steps for all trees. Increases in C allocation to leaves were detected in elevated CO~2~ treatments, while the effects of a 4 month drought were negligible on C allocation in aboveground tissues.  These results reveal how climate change factors impact the investment of photosynthetic C in a *Eucalyptus* tree species and provide evidence that belowground processes may not be as sensitive to global change factors as previously thought. 

\
In conclusion, this PhD research addressed interrelated questions regarding resource allocation in *Eucalyptus* tree species by linking leaf physiology to whole canopy C gain and allocation of photosynthetic C to whole tree growth. This study confirms that the distribution of photosynthetic resources constrain canopy C uptake, yet within canopy leaf physiology does not follow prevailing optimal theory regarding water use. Results from this work reveal how quantifying the fate of photosynthetic C among tissue pools, beyond biomass production, is imperative to accurately assess the impacts of environmental change on tree productivity. This research offers critical empirical data needed to refine process based models which predict canopy C gain from rates of *A~n~* and forest growth models where C allocation is represented. Ultimately, this work contributes valuable information regarding the physiological and growth responses of *Eucalyptus* tree species essential for reconciling the impacts of resource availability and global climate change on Australian ecosystems and the productivity of *Eucalyptus* plantation forests. 

